Just Add IFC! A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-03-02 paper/retrofit.tex
Finish pass over retrofit.

We consider asynchronous communication primitives that allow different
tasks to exchange messages.
When a task is sending a message using |send|, it also labels that
message with a label |il'| (which must be above the tasks current label |il|).
Now, messages can only be received by a task if its current label is
at least as high as the label of the message.

Specifically, receiving a message using
\[ |recvf ix1 ix2 ie1 ie2| \]
binds the senders task identifier and the message
to local variables |ix1| and |ix2|, respectively, and then executes |ie1|.
Otherwise the message gets removed from the queue by rule \textsc{I-noRecv},
and that tasks continues its execution with |ie2|.
We denote the filtering of the message queue by |filter il Q|,
which is defined as follows.
If |Q| is the empty list |nil|, the
function is simply the identity function, i.e.,
|filter il nil = nil|; otherwise:
\[
|filter il ((il', id, ie) , Q )| = \left\{
\begin{array}{l l}
|(il', id, ie) , filter il Q| & \quad \text{if |il' canFlowTo il|}\\
|filter il Q| & \quad \text{otherwise}
\end{array} \right.
\]
This ensures that tasks cannot receive messages that are more secret
than their current label would allow.

@@

\Red{TODO[ds]:
  - clarify discussion on exceptions
  - fix the toLabled
}
Finally, in Figure~\ref{fig:embedding},
we provide all of the rules responsible for actually carrying out the embedding.
The most important feature of this embedding is that every task maintains its own
copy of the target language global state and evaluation context, thus
enforcing isolation between various tasks.  In more detail:

@@

\begin{itemize}
    \item We extend the values, expressions and evaluation contexts of both languages to allow for embeddings terms in one language to be embedded in the other.
    \item We reinterpret |CE| to be evaluation contexts on task pools, providing definitions for |curlyRedE| and |curlyBlueE|.  These rules only operate on the first task in the task pool, which by convention is the only task executing.
    \item We reinterpret |->|, an operation on a single task, in terms of |.->|, operation on task pools.  The correspondence is simple: a task executes a step and then is rescheduled in the task pool according to schedule policy |alpha|.  Not all schedulers preserve non-interference, and we provide two specific schedules in Figure~\ref{fig:scheduler}.
    \item Finally, we define some rules for scheduling, handling sandboxing tasks (which interact with the state of the target language),
    primitives for asynchronously communicating between
    tasks and intermediating between the borders of the two languages.
\end{itemize}

@@

@@

To get a feeling for how these operational semantics operate, we'd like to
describe a few sample programs, and what their outputs are.

\paragraph{Sending mutable references}  As each task maintains its own
global state for the target language, one might reasonably wonder what would
happen if you attempted to send a reference to another task.

\begin{align*}
    & |let i = TI (sandbox (brecvf x i' (IT (!(TI x)))))| \\
    & |in TI (send (IT (ref true)))|
\end{align*}

It would be a violation of information flow if such an operation actually
resulted in references being shared between two tasks.  In fact, the address
will be received by the sandboxed task, but it will be a dangling pointer; the
task will become stuck and removed from execution by \textsc{I-noStep}.  (Readers
who are curious whether or not a static discipline could be enforced to
prevent these dangling pointers should check Section~\ref{sec:extensions}.)

\paragraph{Status of the heap after a sandbox}  Here is a program whose
semantics depends on the definition of $\kappa$:

\begin{align*}
    & |let x = ref true| \\
    & |in TI (sandbox (IT (!x; !x))); x := false|
\end{align*}

When $\kappa$ is the identity, the sandboxed task successfully dereferences
the contents of the reference to be true twice (recall that references
are cloned into the new global state).  When $\kappa$ is the constant function
to the empty store, $x$ becomes a dangling pionter.

\paragraph{Exceptions}  One language feature that has caused difficulty
for many information flow control systems is handling
exceptions.~\cite{Hritcu:2013:YIB:2497621.2498098}  The key difficulty
appears when an exception must cross a stack boundary where the current
label reverts back, i.e. across a \textbf{toLabeled} operator.~\cite{lio}
If the exception crosses this boundary, it can leak information about the
inner sensitive computation.

At first glance, it does not appear that our system has any such operator;
there are only tasks, and the correct behavior for a task throwing
an exception is self-evident: it should simply be removed from the scheduler.
However, under a \emph{sequential scheduler}, the behavior of |sandbox|
can be used to model the behavior of \textbf{toLabeled}:

\begin{align*}
    & |toLabeled l te @= | \\
    & |tab let i = taskId; | \\
    & |tab let j = TI (sandbox (IT (TI (setLabel l); TI (send i l (IT te)))))| \\
    & |tab in (\ _ . TI (setLabel l); TI (recvf x _ x unit))| \\
\end{align*}

When the sequential scheduler is used, the sandboxed task runs to
completion before returning to the task that sandboxed it.  Thus,
a task pool should rather be interpreted as a number of information-flow
aware stack frames.  To read the result of a |toLabeled| operation
(which is communicated via message passing), the parent task must raise
its label; otherwise the label will get dropped, as per the \textsc{I-recv} rule.\footnote{A keen reader may note that with a single, per-task
channel, we might mix up our messages.  Indeed, a proper development requires
the ability to allocate fresh channels; for purposes of simplicity, this
has been omitted.}

In this setting, it is also clear that exceptions should not cross boundaries,
as seen in |toLabeled l fail; true|, which evaluates to |true| no matter even in
the event of an exception inside of the |toLabeled| block:

>>>>

We consider asynchronous communication primitives that allow different
tasks to exchange messages.
When a task is sending a message using |send|, it also labels that
message with a label |il'| (which must be above the tasks current label |il|).
Messages can only be received by a task if its current label is
at least as high as the label of the message.
Specifically, receiving a message using
\[ |recvf ix1 ix2 ie1 ie2| \]
binds the message and the senders task identifier
to local variables |ix1| and |ix2|, respectively, and then executes |ie1|.
Otherwise the message gets removed from the queue by rule \textsc{I-noRecv},
and that tasks continues its execution with |ie2|.
We denote the filtering of the message queue by |filter il Q|,
which is defined as follows.
If |Q| is the empty list |nil|, the
function is simply the identity function, i.e.,
|filter il nil = nil|; otherwise:
\[
|filter il ((il', id, ie) , Q )| = \left\{
\begin{array}{l l}
|(il', id, ie) , filter il Q| & \quad \text{if |il' canFlowTo il|}\\
|filter il Q| & \quad \text{otherwise}
\end{array} \right.
\]
This ensures that tasks cannot receive messages that are more secret
than their current label would allow.

@@

\Red{TODO[ds]:
  - clarify discussion on exceptions
  - fix the toLabled
}
Finally, in Figure~\ref{fig:embedding},
we provide all of the rules responsible for actually carrying out the embedding of the IFC language with the target language.
The most important feature of this embedding is that every task maintains its own
copy of the target language global state and evaluation context, thus
enforcing isolation between various tasks.  In more detail:

@@

\begin{itemize}
    \item We extend the values, expressions and evaluation contexts of both languages to allow for embeddings terms in one language to be embedded in the other.
    \item We reinterpret |CE| to be evaluation contexts on task pools, providing definitions for |curlyRedE| and |curlyBlueE|.  These rules only operate on the first task in the task pool, which by convention is the only task executing.
    \item We reinterpret |->|, an operation on a single task, in terms of |.->|, operation on task pools.  The correspondence is simple: a task executes a step and then is rescheduled in the task pool according to schedule policy |alpha|.  Not all schedulers preserve non-interference, and we provide two specific schedules in Figure~\ref{fig:scheduler}.
    \item Finally, we define some rules for scheduling, handling sandboxing tasks (which interact with the state of the target language),
    and intermediating between the borders of the two languages.
\end{itemize}

@@

We note that it is possible to add rules for a natuaral embedding of
expressions that occur in both languages.  In our specific example
with |targetLangML|, we can choose to preserve booleans across
language boundaries as follows:
\begin{mathpar}
\inferrule[I-bool]
{ |b member {true, false}|  } {| iniE (IT b) -> iniE b |}
\and
\inferrule[T-bool]
{ |b member {true, false}| } {| intE (TI b) -> intE b |}
\end{mathpar}

@@

To get a feeling for how these operational semantics operate, we
would like to
describe a few sample programs, and what their outputs are.

\paragraph{Sending mutable references}  As each task maintains its own
global state for the target language, one might reasonably wonder what would
happen if you attempted to send a reference to another task.
\begin{align*}
    & |let i = TI (sandbox (brecvf x i' (IT (!(TI x)))))| \\
    & |in TI (send (IT (ref true)))|
\end{align*}
It would be a violation of information flow if such an operation actually
resulted in references being shared between two tasks.  In fact, the address
will be received by the sandboxed task, but it will be a dangling pointer; the
task will become stuck and removed from execution by \textsc{I-noStep}.  Readers
who are curious whether or not a static discipline could be enforced to
prevent these dangling pointers should check Section~\ref{sec:extensions}.

\paragraph{Status of the heap after a sandbox}  Here is a program whose
semantics depends on the definition of |klone|:
\begin{align*}
    & |let x = ref true| \\
    & |in TI (sandbox (IT (!x; !x))); x := false|
\end{align*}
When |klone| is the identity, the sandboxed task successfully dereferences
the contents of the reference to be true twice (recall that references
are cloned into the new global state).  When $\kappa$ is the constant function
to the empty store, $x$ becomes a dangling pionter.

\paragraph{Exceptions}  One language feature that has caused difficulty
for many information flow control systems is handling
exceptions~\cite{Hritcu:2013:YIB:2497621.2498098}.  The key difficulty
appears when an exception must cross a stack boundary where the current
label reverts back, i.e. across a |toLabeled| operator~\cite{lio}.
If the exception crosses this boundary, it can leak information about the
inner sensitive computation.

At first glance, it does not appear that our system has any such operator;
there are only tasks, and the correct behavior for a task throwing
an exception is self-evident: it should simply be removed from the scheduler.
However, under a \emph{sequential scheduler} (such as |seqf|),
the behavior of |sandbox|
can be used to model the behavior of such an operator:
\begin{align*}
    & |toLabeled l te @= | \\
    & |tab let i = taskId; | \\
    & |tab let j = TI (sandbox (IT (TI (setLabel l); TI (send i l (IT te)))))| \\
    & |tab in (\ _ . TI (setLabel l); TI (recvf x _ x unit))|
\end{align*}
When the sequential scheduler is used, the sandboxed task runs to
completion before returning to the task that sandboxed it.  Thus,
a task pool should rather be interpreted as a number of information flow
aware stack frames.  To read the result of a |toLabeled| operation
(which is communicated via message passing), the parent task must raise
its label; otherwise the label will get dropped, as per the \textsc{I-recv} rule.\footnote{A keen reader may note that with a single, per-task
channel, we might mix up our messages.  Indeed, a proper development requires
the ability to allocate fresh channels; for purposes of simplicity, this
has been omitted.}

In this setting, it is also clear that exceptions should not cross boundaries,
as seen in |toLabeled l fail; true|, which evaluates to |true| no matter even in
the event of an exception inside of the |toLabeled| block:
\todo{sh}{leaving this to ds, who wanted to finish the example}
